---
title: "Phylogenetic analysis and Visualization"
format: html
---

-   Creating the tree from mammal phylogeny from dataset in R Bioinformatics Cookbook.

```{r}
#install.packages("ape")

#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("treeio")

#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("ggtree")

#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("SGSeq")
```

-   ggtree is used to visualize evolutionary trees.

```{r}
library(rbioinfcookbook)
library(SGSeq)
library(ape)
library(treeio)
library(ggplot2)
library(ggtree)
```

```{r}
newick_file_path <- fs::path_package("extdata", "mammal_tree.nwk", package = "rbioinfcookbook" )

nexus_file_path <- fs::path_package("extdata", "mammal_tree.nexus", package = "rbioinfcookbook")

newick <- ape::read.tree(newick_file_path)
nexus <- ape::read.nexus(nexus_file_path)
```

-   Newick: creates evolutionary trees in text format.

-   Nexus: also creates evolutionary trees in text format but adds DNA sequences or traits in the tree.

-   Both Newick and Nexus falls under ape package.

```{r}
beast_file_path <- fs::path_package("extdata", "beast_mcc.tree", package = "rbioinfcookbook")

raxml_file_path <- fs::path_package("extdata", "RAxML_bipartitionsBranchLabels.H3", package = "rbioinfcookbook")

beast <- read.beast(beast_file_path)
raxml <- read.raxml(raxml_file_path)
```

-   This is the different package call treeio that has two different formats called BEAST and RAxML that is also used to create evolutionary trees.

-   BEAST: uses Bayesian method to include different rates of evolution.

-   RAxML: used for phylogenetic analysis of large datasets. It uses maximum likelihood which is the statistical technique used to find the best value for the evolutionary trees.

```{r}
class(newick)
```

```{r}
class(nexus)
```

```{r}
class(beast)
```

```{r}
class(raxml)
```

```{r}
beast_phylo <- treeio::as.phylo(beast)
newick_tidytree <- treeio::as.treedata(newick)
```

```{r}
treeio::write.beast(newick_tidytree, file = "mammal_tree.beast")
ape::write.nexus(beast_phylo, file = "beast_mcc.nexus")
```

-   Looking at the tree, it was interesting to see that H. sapiens (human) and P. paniscus (chimpanzee) shares the common ancestor. In general, it was interesting to see that apes and human are related and goes a long way back like having shared ancestor with Hylobates.

```{r}
tree_file <- fs::path_package("extdata", "itol.nwk", package = "rbioinfcookbook")

itol<- ape::read.tree(tree_file)
```

```{r}
ggtree(itol)
```

```{r}
ggtree(itol) +
  geom_tiplab(color = "blue", size = 1)
```

-   Using ggtree to add labels to the evolutionary tree.

```{r}
ggtree(itol, layout = "circular") +
  geom_tiplab(color = "blue", size = 2)
```

-   By adding circular layout in the already existing code above, we made the change in how the evolutionary tree could be read.

```{r}
ggtree(itol, layout = "circular") +
  geom_tiplab(color = "blue", size = 2) +
  geom_strip(13, 14, color = "red", barsize = 1)
```

-   Using the color to highlight the particular clade.

```{r}
ggtree(itol) +
  coord_flip() +
  scale_x_reverse()
```

-   The same evolutionary tree could be invereted and read in the different format.

```{r}
ggtree(itol, layout = "unrooted")
```

```{r}
mammal_file <- fs::path_package("extdata", "mammal_tree.nwk", package = "rbioinfcookbook" )

mammal<- ape::read.tree(mammal_file)
```

```{r}
ggtree(mammal) +
  geom_tiplab(color = "blue", size = 2)
```

```{r}
ggtree(mammal, layout = "circular") +
  geom_tiplab(color = "blue", size = 2)
```

-   ggtree is used to create the tree from mammal phylogeny.

### Conclusion

-   I was able to learn to use ggtree to visualize a phylogeny tree and use different layout to create it. I don't think I have any questions in particular for now.
