---
title: "RNA Seq II"
format: html
---

```{r}
#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("ComplexHeatmap")
```

```{r}
#install.packages("viridisLite")
#install.packages("stringr")
#install.packages("RColorBrewer")
#install.packages("circlize")
```

```{r}
library(ComplexHeatmap)
library(viridisLite)
library(stringr)
library(RColorBrewer)
library(circlize)
library(rbioinfcookbook)
```

```{r}
mat <- log(as.matrix(at_tf_gex[ , 5:55]))
ecotype <- stringr::str_split(colnames(mat), ",", simplify = TRUE)[,1]
part <- stringr::str_split(colnames(mat), ",", simplify = TRUE)[,2]
```

-   The data set we are using is at_tf_gex.

-   stringr is used to split the data by ecotype.

```{r}
data_col_func <- circlize::colorRamp2(seq(0, max(mat), length.out = 6), viridisLite::magma(6))

ecotype_colors <- c(RColorBrewer::brewer.pal(12, "Set3"), RColorBrewer::brewer.pal(5, "Set1"))
names(ecotype_colors) <- unique(ecotype)

part_colors <- RColorBrewer::brewer.pal(3, "Accent")
names(part_colors) <- unique(part)
```

-   Using circlize and viridisLite, different color palettes are created for different types of data for plant's ecotype and parts in the heatmap.

```{r}
top_annot <- HeatmapAnnotation("Ecotype" = ecotype, "Plant Part" = part, col = list("Ecotype" = ecotype_colors, "Plant Part" = part_colors), annotation_name_side = "left")

side_annot <- rowAnnotation(length = anno_points(at_tf_gex$Length, pch = 16, size = unit(1, "mm"), axis_param = list(at = seq(1, max(at_tf_gex$Length), length.out = 4)),))
```

-   top_annot is used to label the information on ecotype and plant part.

-   annotation_name_side: label the left of the color that they represent.

-   side_annot: labels the rows of the heatmap.

-   anno_points(): tells where exactly the points needs to be plotted.

-   pch: tells the shape of the point.

-   size: tells the size of the point.

-   axis_param: tells the location of the ticks on the x-axis.

```{r}
ht_1 <- Heatmap(mat, name="log(TPM)", row_km = 6, col = data_col_func, top_annotation = top_annot, right_annotation = side_annot, cluster_columns = TRUE, column_split = ecotype, show_column_names = FALSE, column_title = " ")

ComplexHeatmap::draw(ht_1)
```

-   row_km: set the number of clusters for the rows.

-   cluster_colums: makes sure that columns in the heatmap are clustered together based on how similar is their gene expression.

-   column_split: groups the column by ecotype.

-   show_column_name: hides the column name to make the heat map look less clustered.

### Conclusion

-   I have learnt how to use R code to be able to filter out and select the required data sets and I think I am becoming more comfortable using R studio.

-   The question I still have is how to read the heat map? It looks pretty confusing.
