---
title: "RNA seq1"
format: html
---

```{r}
#install.packages("devtools")
library(devtools)
```

```{r}
#devtools::install_github("danmaclean/rbioinfcookbook")
library(rbioinfcookbook)
```

```{r}
#install.packages("forcats")

#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("edgeR")

#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("Biobase")

library(forcats)
library(edgeR)
```

```{r}
library(Biobase)

```

```{r}
genes <- count_dataframe[['gene']]
count_dataframe[['gene']] <- NULL
count_matrix <- as.matrix(count_dataframe)
rownames(count_matrix) <- genes
```

```{r}
experiments_of_interest <- c("L1Larvae", "L2Larvae")
columns_of_interest <- which(pheno_data[['stage']] %in% experiments_of_interest)
```

```{r}
grouping <- pheno_data[["stage"]] [columns_of_interest] |> forcats::as_factor()
```

```{r}
counts_of_interest <- count_matrix[,counts = columns_of_interest]
```

```{r}
count_dge <- edgeR::DGEList(counts = counts_of_interest, group = grouping)

```

```{r}
design <- model.matrix(~grouping)
eset_dge <- edgeR::estimateDisp(count_dge, design)
fit <- edgeR::glmQLFit(eset_dge, design)
result <- edgeR::glmQLFTest(fit, coef=2)
topTags(result)
```

All the gene above is of the species D.melanogaster.ss

The name of the gene FBgn0027527 is Osiris 6. It is located in endomembrane system and is expresed in embryonic dorsal epidermis, embryonic head epidermis, embryonic hypopharynx, embryonic ventral epidermis and gut section.

The name of the gene FBgn0243586 is Tbby. It is predicted to be involved in body morphogenesis and chitin-based cuticle development and located in extracellular matrix. It is expressed in dorsal head epidermis; embryonic dorsal epidermis; embryonic ventral epidermis; lateral head epidermis; and ventral head epidermis.

What I learnt is how to use different codes to get the specific RNA transcripts that we need to research from different genes of D.melanogaster using edgeR.
